Suplementary Code 3: Run Simulations on the Last Birth Omited Data

### 2023-12-27 S3_response_to_reviewers.Rmd

---
title: 'Suplementary Material 3: Code to Respond to Reviewers'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))){
 if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx"){
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
} 
}

options(digits = 9)
options(scipen = 9)
options(knitr.kable.NA="")
```

## Introduction

The aim of this document is to respond to the reviewers comments. The comments are:

1) I would like to see whether indeed mothers are statistically more likely to stop having children after a twin birth (perhaps I missed that analysis and it was indeed presented)?

2) Directly test the odds that a mother terminate reproduction after a twin birth? In other words, use statistics to assess post-twinning reproductive outcomes compared to post-singleton reproductive outcomes.

3) Plot the real data as a cloud of points behind these lines- it is important to see the actual dataset in analyses like these.

4) Look at the analyses when limiting to opposite-sex offspring?

5) Based on analyses of Finnish data from 1953-1964 Eriksson and Fellman show that illegitimate maternities have a higher twinning rate than legitimate ones when maternal age and birth order are taken into account. They speculate that unmarried mothers are women who tend to become pregnant more easily than others owing to the effect of certain factors, e.g. a higher ovulation rate and to have relatively better physical qualifications for going through a twin pregnancy than mothers in general. 
It seems not unlikely that such mothers may remain single of the twin pregnancy, especially in earlier times?

```{r}
#cleanup memory
gcstuff <- gc(verbose=FALSE); rm(gcstuff);
```

```{r}
#get last birth adding function
source("./R/last_birth.R")

#simplified twinR summary tables 
source("./R/twinR_summary.R")

#fix twinR compute predictions to do prediction with no lambda  as well
source("./R/twinR_predictions.R")

#simple convenience functions
source("./R/utils.R")
```

```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- min(c(50L, parallel::detectCores() - 1))

## Set option in spaMM:
spaMM::spaMM.options(nb_cores = nb_cores)
```

## Data Import

The Estonian dataset has been formatted to include the same columns as *the data_births_all* dataset from the **twinR** package. The only difference is that the columns *pop* and *monthly* are excluded as these are constant.

```{r}
#some labels
estLab <- "Estonian"
westLab <- "9 other European"
```

```{r}
#Import and preproccess Estonian Data

data_births_monthly_EE <- readRDS("./data/data_births_all_EE.rds")

#the twinR package expects population to be present
data_births_monthly_EE$pop <- "Estonia"

## Expand the birth level data for the fit of statistical models:
data_births_monthly_EE <- twinR::expand_data(data_births_monthly_EE)

data_births_monthly_EE <- add_last_birth(data_births_monthly_EE)

data_births_monthly_EE_not_last <- data_births_monthly_EE[!data_births_monthly_EE$last,]

#make the aggregates
dmm_EE <- twinR::aggregate_data(data_births_monthly_EE)
dmm_EE$prob_twin <- dmm_EE$twin_total / dmm_EE$births_total

dmm_EE_nl <- twinR::aggregate_data(data_births_monthly_EE_not_last)
dmm_EE_nl$prob_twin <- dmm_EE_nl$twin_total / dmm_EE_nl$births_total
```

```{r}
##Import and pre-proccess twinR package data

## Filter the raw data to only keep data with monthly resolution:
data_births_monthly <- twinR::filter_data(twinR::data_births_all)

## Expand the birth level data for the fit of statistical models:
data_births_monthly <- twinR::expand_data(data_births_monthly)

data_births_monthly <- add_last_birth(data_births_monthly)
data_births_monthly_not_last <- data_births_monthly[!data_births_monthly$last,]

dmm_orig <- twinR::aggregate_data(data_births_monthly)
dmm_orig$prob_twin <- dmm_orig$twin_total /dmm_orig$births_total

dmm_orig_nl <- twinR::aggregate_data(data_births_monthly_not_last)
dmm_orig_nl$prob_twin <- dmm_orig_nl$twin_total /dmm_orig_nl$births_total

```

```{r}
#compared to twinR original add data on :
#-age at first birth (AFB)
#-quantiles range mother birth year
#-quantiles range offspring birth year
#-mean + SE total births
all_tbls <- rbind(build_data_summary.table(data_births_monthly),
                  build_data_summary.table(data_births_monthly_not_last),
                  build_data_summary.table(data_births_monthly_EE),
                  build_data_summary.table(data_births_monthly_EE_not_last))
```

## Fitting models

### Comment 1: Logistic regression

We can test the hypothesis that mothers are statistically more likely to stop having children after a twin birth by fitting a logistic regression model with the response variable being the binary variable *last* and the predictor variable being the binary variable *twin*.

```{r}
#fit logistic regression Estonia
logit_fit_EE <- glm(last ~ twin,
                 data = data_births_monthly_EE,
                 family = binomial(link = "logit"))
summary(logit_fit_EE)
```

```{r}
#fit logistic regression Other European
logit_fit <- glm(last ~ twin,
                 data = data_births_monthly,
                 family = binomial(link = "logit"))
summary(logit_fit)
```
### Comment 2: Proportion of twins

The same test gives us the odds of mothers having more children after a twin birth.

```{r}
# Extract coefficients
coef <- coef(logit_fit_EE)

# Calculate odds ratios
odds_ratios <- exp(coef)
```
For the Estonian dataset the  odds of a birth event being the last one for a mother are `r round(odds_ratios[2], 3)` times higher for a twin birth than for a singleton birth.

```{r}
# Extract coefficients
coef <- coef(logit_fit)

# Calculate odds ratios
odds_ratios <- exp(coef)

```
For the other dataset the odds of a birth event being the last one for a mother are `r round(odds_ratios[2], 3)` times higher for a twin birth than for a singleton birth.

## Comment 3: Original Data on Plots

The reviewers also asked to put the data on the Figure 1.

```{r}
# import the function to do model fit and predictions
source("./R/fit_models.R")
```

```{r}
## full data

## Estonia - mother level data
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_fit <- fitPredictions(dmm_EE, formula, predict = T)

## TwinR - mother level data
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + (1|pop)"
dmm_orig_fit <- fitPredictions(dmm_orig, formula, predict=T)

```

```{r}
library(ggplot2)
#some nice colors
bc <- c("azure4","purple4", "black", "navy", "darkgoldenrod2", "springgreen3", "gray")

#use new base theme that displays also grid lines
source("./R/twinR_theme.R")
```

### Fig 1a: Estonian vs TwinR Full Data

```{r}

fig2_EE_plot_data <- dmm_EE_fit$results
fig2_orig_plot_data <- dmm_orig_fit$results
```

```{r, fig.height=11.7/2, fig.width=8.3}
infoTxt <- "model prediction with 95% CI and data with values"

fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color=estLab),  size = 1) +
  geom_point(data=dmm_EE[dmm_EE$births_total <17, ],
               aes(x=births_total, y=prob_twin, color=estLab, fill = estLab),
               alpha=0.1,
               position = position_nudge(x = -0.1)) +
    stat_summary(data=dmm_EE[dmm_EE$births_total <17, ],
                 aes(x=births_total, y=prob_twin, color=estLab, fill = estLab),
                 alpha=0.1,
                 position = position_jitter(),
                 fun.data=mean_se) +

    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color=estLab, fill = estLab),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data,
             aes(y = estimate,  x=births_total, color=westLab),  size = 1) +
  geom_point(data=dmm_orig[dmm_orig$births_total <19, ],
               aes(x=births_total, y=prob_twin, color=estLab, fill = estLab),
               alpha=0.1,
               position = position_jitter()) +
    stat_summary(data=dmm_orig[dmm_orig$births_total <19, ],
                 aes(x=births_total, y=prob_twin,
                     color=westLab, fill=westLab),
                 alpha=1,
                 shape = 1,
                 position = position_nudge(x = 0.1),
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_orig_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill=westLab),
                alpha = 0.1) +
    ggplot2::scale_x_continuous(breaks = 1:18) +

    ggplot2::coord_cartesian() +
    labs(subtitle = paste0("(a) All births, ", infoTxt),
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "population")

p2
```

The first plot clearly illustrates that we cannot plot the whole data itself on the plot, as there are women who only gave birth to twins. We can plot a subset (eg those with at least 10 children) but this would lose the point of the plot. There is just so much data that is semi-categorical that we cannot think of a way put into a plot so that it gives additional information.

## Comment 4: Dizygotic Twins

## Comment 5: Unmarried Mothers

```{r}
#END
```

### 2023-12-27 S3_response_to_reviewers.Rmd

```{r}
# Extract coefficients
coef <- coef(logit_fit)

# Calculate odds ratios
odds_ratios <- exp(coef)

```
For the other dataset the odds of a birth event being the last one for a mother are `r round(odds_ratios[2], 3)` times higher for a twin birth than for a singleton birth.

```{r}
#compare the precentage of twin births between last births and not last births
cat("Estonian dataset:\n")
tbl <- table(data_births_monthly_EE[c("twin", "last")])
tbl
cat("\n")
tbl <- tbl / sum(tbl)
#now calcualte the % by hand
cat("Last twins: ", round(100 * tbl[2,2] / sum(tbl[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tbl[2,1] / sum(tbl[,1]), 2), "%\n")

```{r}
#compare the precentage of twin births between last births and not last births
cat("Western dataset:\n")
tbl <- table(data_births_monthly[c("twin", "last")])
tbl
cat("\n")
tbl <- tbl / sum(tbl)
#now calcualte the % by hand
cat("Last twins: ", round(100 * tbl[2,2] / sum(tbl[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tbl[2,1] / sum(tbl[,1]), 2), "%\n")

### 2023-12-27 S3_response_to_reviewers.Rmd

```{r}
# Extract coefficients
coef <- coef(logit_fit)

# Calculate odds ratios
odds_ratios <- exp(coef)

```
For the other dataset the odds of a birth event being the last one for a mother are `r round(odds_ratios[2], 3)` times higher for a twin birth than for a singleton birth.

```{r}
#compare the precentage of twin births between last births and not last births
cat("Estonian dataset:\n")
tbl <- table(data_births_monthly_EE[c("twin", "last")])
tbl
cat("\n")
#now calcualte the % by hand
cat("Last twins: ", round(100 * tbl[2,2] / sum(tbl[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tbl[2,1] / sum(tbl[,1]), 2), "%\n")

```{r}
#compare the precentage of twin births between last births and not last births
cat("Western dataset:\n")
tbl <- table(data_births_monthly[c("twin", "last")])
tbl
cat("\n")
#now calcualte the % by hand
cat("Last twins: ", round(100 * tbl[2,2] / sum(tbl[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tbl[2,1] / sum(tbl[,1]), 2), "%\n")

### 2024-01-03 S5_response_to_reviewers.Rmd

---
title: 'Suplementary Material 3: Code to Respond to Reviewers'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

@@

### Comment 1: Logistic regression

We can test the hypothesis that mothers are statistically more likely to stop having children after a twin birth by fitting a logistic regression model with the response variable being the binary variable *last* and the predictor variable being the binary variable *twin*.

```{r}
#fit logistic regression Estonia
logit_fit_EE <- glm(last ~ twin,
                 data = data_births_monthly_EE,
                 family = binomial(link = "logit"))
summary(logit_fit_EE)
```

```{r}
#fit logistic regression Other European
logit_fit <- glm(last ~ twin,
                 data = data_births_monthly,
                 family = binomial(link = "logit"))
summary(logit_fit)
```

### Comment 2: Proportion of twins

@@

```{r}
# Extract coefficients
coef <- coef(logit_fit)

# Calculate odds ratios
odds_ratios <- exp(coef)

```
For the other dataset the odds of a birth event being the last one for a mother are `r round(odds_ratios[2], 3)` times higher for a twin birth than for a singleton birth.

```{r}
#compare the precentage of twin births between last births and not last births
cat("Estonian dataset:\n")
tble <- table(data_births_monthly_EE[c("twin", "last")])
tble
cat("\n")
#now calcualte the % by hand
cat("Last twins: ", round(100 * tble[2,2] / sum(tble[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tble[2,1] / sum(tble[,1]), 2), "%\n")

@@

```

```{r}
#compare the precentage of twin births between last births and not last births
cat("Western dataset:\n")
tblw <- table(data_births_monthly[c("twin", "last")])
tblw
cat("\n")
#now calcualte the % by hand
cat("Last twins: ", round(100 * tblw[2,2] / sum(tblw[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tblw[2,1] / sum(tblw[,1]), 2), "%\n")

@@

```

```{r}
# Combine the two tables into a 3-dimensional array
tbl <- array(c(tble, tblw), dim = c(dim(tble), 2))

#numeric problem workaround
tbl <- tbl / 1

# Perform the Mantel-Haenszel test
mantelhaen.test(tbl)
```

@@

    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color=estLab, fill = estLab),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data,
             aes(y = estimate,  x=births_total, color=westLab),  size = 1) +
  geom_point(data=dmm_orig[dmm_orig$births_total <19, ],
               aes(x=births_total, y=prob_twin, color=estLab, fill = estLab),
               alpha=0.1,
               position = position_jitter()) +
    stat_summary(data=dmm_orig[dmm_orig$births_total <19, ],
                 aes(x=births_total, y=prob_twin,
                     color=westLab, fill=westLab),
                 alpha=1,
                 shape = 1,
                 position = position_nudge(x = 0.1),
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_orig_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill=westLab),
                alpha = 0.1) +
    ggplot2::scale_x_continuous(breaks = 1:18) +

    ggplot2::coord_cartesian() +
    labs(subtitle = paste0("(a) All births, ", infoTxt),
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "population")

p2
```

The first plot clearly illustrates that we cannot plot the whole data itself on the plot, as there are women who only gave birth to twins. We can plot a subset (e.g. those with at least 10 children) but this would lose the point of the plot. There is just so much data that is semi-categorical that we cannot think of a way put into a plot so that it gives additional information.

@@

## Comment 4: Dizygotic Twins

We can do the same analysis for dizygotic twins.

```{r}
#re-import the data
dbm_EE <- readRDS("./data/data_births_all_EE.rds")
```

```{r}
#take out mothers that had twins of single sex
single_sex_twin_mothers <- as.character(unlist(unique(
  dbm_EE[dbm_EE$twin & dbm_EE$cf_sex != "mixed", "maternal_id"])
  ))

dbmEE_wo_single_sex_twins <- dbm_EE[!(dbm_EE$maternal_id %in% 
                                        single_sex_twin_mothers),]
```

```{r}
#the twinR package expects population to be present
dbmEE_wo_single_sex_twins$pop <- "Estonia"

## Expand the birth level data for the fit of statistical models:
dbmEE_wo_single_sex_twins <- twinR::expand_data(dbmEE_wo_single_sex_twins)

dbmEE_wo_single_sex_twins <- add_last_birth(dbmEE_wo_single_sex_twins)

dbmEE_wo_single_sex_twins_not_last <- dbmEE_wo_single_sex_twins[!dbmEE_wo_single_sex_twins$last,]

#make the aggregates
dmm_EE_mixed <- twinR::aggregate_data(dbmEE_wo_single_sex_twins)
dmm_EE_mixed$prob_twin <- dmm_EE_mixed$twin_total / dmm_EE_mixed$births_total

dmm_EE_mixed_nl <- twinR::aggregate_data(dbmEE_wo_single_sex_twins_not_last)
dmm_EE_mixed_nl$prob_twin <- dmm_EE_mixed_nl$twin_total / dmm_EE_mixed_nl$births_total
```

### Full Data

```{r}
options(scipen = 99)
```

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_mixed_fit <- fitPredictions(dmm_EE_mixed, formula)

knitr::kable(build_fit_summary.table(dmm_EE_mixed_fit$fit))
```

### No Last Births Data

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_mixed_nl_fit <- fitPredictions(dmm_EE_mixed_nl, formula)

knitr::kable(build_fit_summary.table(dmm_EE_mixed_nl_fit$fit))
```

### Plot the Predictions

```{r}
fig2_EE_plot_data <- dmm_EE_mixed_fit$results
fig2_EE_plot_data_nl <- dmm_EE_mixed_nl_fit$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
```

```{r}
dmm_EE_mixed_nl_plot <- dmm_EE_mixed_nl
dmm_EE_mixed_nl_plot$births_total <- dmm_EE_mixed_nl_plot$births_total + 1
```

```{r}
#colors
bc <- c("purple", "black")
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color="EE all"),
            linewidth = 1) +
    stat_summary(data=dmm_EE_mixed[dmm_EE_mixed$births_total <14, ],
                 aes(x=births_total, y=prob_twin, color="EE all",
                     fill = "EE all"), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total,
                    ymin = lwr, ymax = upr, color="EE all", fill = "EE all"),
                alpha = 0.3) + 
    geom_line(data=fig2_EE_plot_data_nl,
              aes(y = estimate,  x=births_total, color="EE w/o last"),
              linewidth = 1) +
    stat_summary(data=dmm_EE_mixed_nl_plot[dmm_EE_mixed_nl_plot$births_total <14,],
                 aes(x=births_total, y=prob_twin, color="EE w/o last",
                     fill = "EE w/o last"), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data_nl,
                aes(y = estimate,  x=births_total, ymin = lwr,
                    ymax = upr, color="EE w/o last", fill = "EE w/o last"),
                alpha = 0.3) + 
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.002)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Prediction only mixed sex twins + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "dataset")

The mixed sex twinning rate seem to remain constant if last births are excluded from the data and slightly decline if last births are included. The decline seems not to be significant.

## Comment 5: Illegitimate Maternities

We can  add a predictor to the model describing if the women had first birth or first conception before being married. We will do both to investigate if illegitimate maternities affect twinning rate.

```{r}
#add the illigimate children columns
illigimate_birth <- unique(data_births_monthly_EE[c( "illigimate_birth",
                                                     "maternal_id")])
illigimate_birth <- setNames(illigimate_birth$illigimate_birth,
                             illigimate_birth$maternal_id)

illigimate_conception <- unique(data_births_monthly_EE[c( "illigimate_conception","maternal_id")])
illigimate_conception <- setNames(illigimate_conception$illigimate_conception, illigimate_conception$maternal_id)
```

```{r}
dmm_EE$illigimate_conception <- illigimate_conception[dmm_EE$maternal_id]

dmm_EE$illigimate_birth <- illigimate_birth[dmm_EE$maternal_id]
```

```{r}
dmm_EE_nl$illigimate_conception <- illigimate_conception[dmm_EE_nl$maternal_id]

dmm_EE_nl$illigimate_birth <- illigimate_birth[dmm_EE_nl$maternal_id]
```

```{r}
cat("First birth before marriage:\n")
table(dmm_EE$illigimate_birth)
```

```{r}
cat("First conception before marriage:\n")
table(dmm_EE$illigimate_conception)
```

### With All Data

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + illigimate_birth"
dmm_EE_fit_birth <- fitPredictions(dmm_EE, formula)

knitr::kable(build_fit_summary.table(dmm_EE_fit_birth$fit))
```

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + illigimate_conception"
dmm_EE_fit_conception <- fitPredictions(dmm_EE, formula)

knitr::kable(build_fit_summary.table(dmm_EE_fit_conception$fit))
```

### No Last Births Data

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + illigimate_birth"
dmm_EE_nl_fit_birth <- fitPredictions(dmm_EE_nl, formula)

knitr::kable(build_fit_summary.table(dmm_EE_nl_fit_birth$fit))
```

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + illigimate_conception"
dmm_EE_fit_nl_conception <- fitPredictions(dmm_EE_nl, formula)

knitr::kable(build_fit_summary.table(dmm_EE_fit_nl_conception$fit))
```

### Figures for Estonian mothers with and without Last Birth

#### Illigimate Birth

```{r}
fig2_EE_plot_data <- dmm_EE_fit_birth$results
fig2_EE_plot_data_nl <- dmm_EE_nl_fit_birth$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
```

```{r}
dmm_EE_nl_plot <- dmm_EE_nl
dmm_EE_nl_plot$births_total <- dmm_EE_nl_plot$births_total + 1
```

```{r, fig.width=6, fig.height=4}
#colors
bc <- c("purple", "black")
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color=illigimate_birth),
            linewidth = 1) +
    stat_summary(data=dmm_EE[dmm_EE$births_total <13, ],
                 aes(x=births_total, y=prob_twin, color=illigimate_birth,
                     fill = illigimate_birth), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color= illigimate_birth, fill = illigimate_birth),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Prediction from full data model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I birth")
cat("\n")
p2 
cat("\n")
```

```{r}
#colors
bc <- c("purple", "black")
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data_nl,
              aes(y = estimate,  x=births_total, color=illigimate_birth), 
            linewidth = 1) +
    stat_summary(data=dmm_EE_nl[dmm_EE_nl$births_total <12, ],
                 aes(x=births_total, y=prob_twin, color=illigimate_birth,
                     fill = illigimate_birth), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data_nl,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color= illigimate_birth, fill = illigimate_birth),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Prediction from w/o last model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I birth")
cat("\n")
p2 
cat("\n")
```

#### Illigimate conception

```{r}
fig2_EE_plot_data <- dmm_EE_fit_conception$results
fig2_EE_plot_data_nl <- dmm_EE_fit_nl_conception$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
```

```{r}
dmm_EE_nl_plot <- dmm_EE_nl
dmm_EE_nl_plot$births_total <- dmm_EE_nl_plot$births_total + 1
```

```{r, fig.width=6, fig.height=4}
#colors
bc <- c("purple", "black")
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color=illigimate_conception),
            linewidth = 1) +
    stat_summary(data=dmm_EE[dmm_EE$births_total <13, ],
                 aes(x=births_total, y=prob_twin, color=illigimate_conception,
                     fill = illigimate_conception), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color= illigimate_conception, fill = illigimate_conception),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Prediction from full data model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I conception")
cat("\n")
p2 
cat("\n")
```

```{r}
#colors
bc <- c("purple", "black")
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data_nl,
              aes(y = estimate,  x=births_total, color=illigimate_conception),
            linewidth = 1) +
    stat_summary(data=dmm_EE_nl[dmm_EE_nl$births_total <12, ],
                 aes(x=births_total, y=prob_twin, color=illigimate_conception,
                     fill = illigimate_conception), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data_nl,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color= illigimate_conception, fill = illigimate_conception),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Prediction from w/o last model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + 
  labs(fill = "Illigimate I conception")
cat("\n")
p2
cat("\n")
```

### 2024-01-03 S5_response_to_reviewers.Rmd

knitr::kable(build_fit_summary.table(dmm_EE_nl_fit_birth$fit))
```

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + illigimate_conception"
dmm_EE_fit_nl_conception <- fitPredictions(dmm_EE_nl, formula)

knitr::kable(build_fit_summary.table(dmm_EE_fit_nl_conception$fit))
```

### Illigimate Birth Figures for Estonian mothers With and Without Last Birth

```{r}
#' Plot Binary Predictions
#'
#' This function generates a ggplot based on the provided parameters.
#'
#' @param plot_data The data to be plotted.
#' @param summary_data The data to be summarized.
#' @param color_var The color variable for the plot.
#' @param threshold The threshold for births_total mean data to display.
#'
#' @return A ggplot object.
#'
plot_predictions <- function(plot_data, summary_data, color_var, threshold = 13) {
  fig2_ext_orig <- ggplot() +  
    geom_line(data=plot_data,
              aes_string(y = "estimate",  x="births_total", color=color_var),
              linewidth = 1) +
    stat_summary(data=summary_data[summary_data$births_total < threshold, ],
                 aes_string(x="births_total", y="prob_twin", color=color_var,
                            fill = color_var), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=plot_data,
                aes_string(y = "estimate",  x="births_total", ymin = "lwr",
                           ymax = "upr",
                           color=color_var, fill = color_var),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian()
  return(fig2_ext_orig)
}

@@

```

```{r}
fig2_EE_plot_data <- dmm_EE_fit_birth$results
fig2_EE_plot_data_nl <- dmm_EE_nl_fit_birth$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
```

```{r}
dmm_EE_nl_plot <- dmm_EE_nl
dmm_EE_nl_plot$births_total <- dmm_EE_nl_plot$births_total + 1
```

```{r}
plot_predictions(plot_data = fig2_EE_plot_data,
                 summary_data = dmm_EE,
                 color_var = "illigimate_birth") +
  labs(subtitle = "Prediction from full data model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births") +
  base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I birth")
```

```{r}
plot_predictions(plot_data = fig2_EE_plot_data_nl,
                 summary_data = dmm_EE_nl_plot,
                 color_var = "illigimate_birth") +
    labs(subtitle = "Prediction from w/o last model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")+ 
  base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I birth")

### Illigimate Conception Figures for Estonian mothers With and Without Last Birth

@@

### Illigimate Conception Figures for Estonian mothers With and Without Last Birth

```{r}
fig2_EE_plot_data <- dmm_EE_fit_conception$results
fig2_EE_plot_data_nl <- dmm_EE_fit_nl_conception$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
```

```{r}
plot_predictions(plot_data = fig2_EE_plot_data,
                 summary_data = dmm_EE,
                 color_var = "illigimate_conception",
                 threshold = 12) +
  labs(subtitle = "Prediction from full data model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births") +
  base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I conception")
```

```{r}
plot_predictions(plot_data = fig2_EE_plot_data_nl,
                 summary_data = dmm_EE_nl_plot,
                 color_var = "illigimate_conception",
                 threshold = 12) +
    labs(subtitle = "Prediction from w/o last model + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")+ 
  base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") +
  labs(fill = "Illigimate I conception")

### 2024-01-04 S1_replicate_stats_and_plots.Rmd
rename title

---
title: 'Suplementary Material 1: Code to Reproduce Analysis and Plots'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

>>>>

---
title: 'Suplementary Code 1: Code to Reproduce Analysis and Plots'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

### 2024-01-04 S2_draw_publication_plots.Rmd
rename title

---
title: 'Suplementary Material 2: Drawing Publication Plots'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

>>>>

---
title: 'Suplementary Code 2: Drawing Publication Plots'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

### 2024-01-04 S3_run_simulations_EE_not_last.Rmd
rename title

---
title: 'Suplementary Material 3: Run Simulations on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

>>>>

---
title: 'Suplementary Code 3: Run Simulations on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---
